---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

In SQL you can filter rows by using WHERE and columns by using SELECT. Let's look at an example by using the Lahman baseball database. First, we ahould load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at some examples. 

Suppose we want to see the homerun totals for the 1927 Yankees. We could write:

```{r}
query<-"select playerID,teamID,yearID,HR from Batting
where teamID='NYA' and yearID=1927"
sqldf(query)
```
 Suppose we want to find all insances of Phillies in the 1970s hitting more than 30 homeruns.

```{r}
query<-"select playerID,teamID,yearID,HR from Batting
where teamID='PHI' and HR>30 and yearID>=1970 and yearID<1980"

sqldf(query)
```

SUppose we want to find all instances of Yankees hitting 40 homeruns or more.

```{r}
query<-"select playerID,teamID,yearID,HR from Batting where teamID='NYA' and HR>=40"

sqldf(query)
```

Suppose we want to find all the instances where a player had more than 40 homeruns but less than 60 strikeouts.

```{r}
query<-"select playerID,teamID,yearID,HR,SO from Batting
where HR>40 and SO<60"

sqldf(query)

```

### Order By

With SQL we can also order the rows in the output. For example; suppose we want to see every instance of a playr hitting more than 50 HR but youd like the player with the most HR to be at the top. Do the following:

```{r}
query<-"select playerID,yearID,teamID,HR from Batting 
where HR>50
order by HR desc"

sqldf(query)
```

### Aggregation
With SQL we can squash chunks of data together to get things such as sumations,averages,maximuns, and minimums. For example; Suppose we want to find what players have the highest average per season Hoomerun totals, Listing only those with averages greater than 30.

```{r}
query<-" select playerID, avg(HR) from Batting
group by playerID
having avg(HR)>30
order by avg(HR) desc"

sqldf(query)
```

